### 2025-04-17 multiple-regression.qmd
20250417 - plots

```{r}
#| label: fig-distributionQQplots
#| fig-cap: "Quantile–Quantile (QQ) Plots of Various Distributions."
#| fig-alt: "Quantile–Quantile (QQ) Plots of Various Distributions."
#| fig-subcap:
#|   - "Normal Distribution"
#|   - "Normal Distribution"
#|   - "Bimodal Distribution"
#|   - "Bimodal Distribution"
#|   - "Negatively Skewed Distribution"
#|   - "Negatively Skewed Distribution"
#|   - "Positively Skewed Distribution"
#|   - "Positively Skewed Distribution"
#|   - "Platykurtic (Light-Tailed) Distribution"
#|   - "Platykurtic (Light-Tailed) Distribution"
#|   - "Lepokurtic (Heavy-Tailed) Distribution"
#|   - "Lepokurtic (Heavy-Tailed) Distribution"
#| code-fold: true

@@

hist(
  distribution_normal,
  main = "Normal Distribution",
  col = "skyblue")

@@

hist(
  distribution_bimodal,
  main = "Bimodal",
  col = "skyblue")

@@

hist(
  distribution_negativeSkew,
  main = "Negatively Skewed",
  col = "skyblue")

@@

hist(
  distribution_positiveSkew,
  main = "Positively Skewed",
  col = "skyblue")

@@

hist(
  distribution_lightTailed,
  main = "Platykurtic (Light Tailed)",
  col = "skyblue")

@@

hist(
  distribution_heavyTailed,
  main = "Leptokurtic (Heavy Tailed)",
  col = "skyblue")

@@

To address multicollinearity, you can drop a redundant predictor or you can also use [principal component analysis](#sec-pca) or [factor analysis](#sec-factorAnalysis) of the [predictors](#sec-correlationalStudy) to reduce the [predictors](#sec-correlationalStudy) down to a smaller number of meaningful [predictors](#sec-correlationalStudy).
For a meaningful answer in a regression framework that is precise and confident, you need a low level of intercorrelation among [predictors](#sec-correlationalStudy), unless you have a very large sample size.

@@

Multiple regression allows examining the association between multiple [predictor variables](#sec-correlationalStudy) and one [outcome variable](#sec-correlationalStudy).
Inclusion of multiple [predictors](#sec-correlationalStudy) in the model allows for potentially greater predictive accuracy and identification of the extent to which each variable uniquely contributes to the [outcome variable](#sec-correlationalStudy).
As with [correlation](#sec-correlation), an association does not imply causation.
However, identifying associations is important because associations are a necessary (but insufficient) condition for causality.
When developing a multiple regression model, it is important to pay attention for potential [multicollinearity](#sec-multipleRegressionMulticollinearity)—it may become difficult to detect a given [predictor variable](#sec-correlationalStudy) as [statistically significant](#sec-statisticalSignificance) due to the greater uncertainty around the parameter estimates.

>>>>

```{r}
#| layout-ncol: 2
#| label: fig-distributionQQplots
#| fig-cap: "Quantile–Quantile (QQ) Plots of Various Distributions."
#| fig-alt: "Quantile–Quantile (QQ) Plots of Various Distributions."
#| fig-subcap:
#|   - "Normal Distribution"
#|   - "Normal Distribution"
#|   - "Bimodal Distribution"
#|   - "Bimodal Distribution"
#|   - "Negatively Skewed Distribution"
#|   - "Negatively Skewed Distribution"
#|   - "Positively Skewed Distribution"
#|   - "Positively Skewed Distribution"
#|   - "Platykurtic (Light-Tailed) Distribution"
#|   - "Platykurtic (Light-Tailed) Distribution"
#|   - "Leptokurtic (Heavy-Tailed) Distribution"
#|   - "Leptokurtic (Heavy-Tailed) Distribution"
#| code-fold: true

@@

hist(
  distribution_normal,
  main = "Normal Distribution",
  col = "#0099F8")

@@

hist(
  distribution_bimodal,
  main = "Bimodal Distribution",
  col = "#0099F8")

@@

hist(
  distribution_negativeSkew,
  main = "Negatively Skewed Distribution",
  col = "#0099F8")

@@

hist(
  distribution_positiveSkew,
  main = "Positively Skewed Distribution",
  col = "#0099F8")

@@

hist(
  distribution_lightTailed,
  main = "Platykurtic (Light Tailed) Distribution",
  col = "#0099F8")

@@

hist(
  distribution_heavyTailed,
  main = "Leptokurtic (Heavy Tailed) Distribution",
  col = "#0099F8")

@@

To address multicollinearity, you can drop a redundant predictor or you can also use [principal component analysis](#sec-pca) or [factor analysis](#sec-factorAnalysis) of the [predictors](#sec-correlationalStudy) to reduce the [predictors](#sec-correlationalStudy) down to a smaller number of meaningful [predictors](#sec-correlationalStudy).
For a meaningful answer regarding [predictors](#sec-correlationalStudy) in a regression framework that is precise and confident, you need a low level of intercorrelation among [predictors](#sec-correlationalStudy), unless you have a very large sample size.
However, if you are merely interested in prediction—and are not interested in interpreting the regression coefficients of individual predictors—multicollinearity poses less of a problem.
For instance, [machine learning](#sec-machineLearning) cares more about achieving the greatest predictive accuracy possible and cares less about explaining which [predictors](#sec-correlationalStudy) are causally related to the [outcome](#sec-correlationalStudy).
So, multicollinearity is less of a concern for [machine learning](#sec-machineLearning) approaches.

@@

Multiple regression allows examining the association between multiple [predictor variables](#sec-correlationalStudy) and one [outcome variable](#sec-correlationalStudy).
Inclusion of multiple [predictors](#sec-correlationalStudy) in the model allows for potentially greater predictive accuracy and identification of the extent to which each variable uniquely contributes to the [outcome variable](#sec-correlationalStudy).
As with [correlation](#sec-correlation), an association does not imply causation.
However, identifying associations is important because associations are a necessary (but insufficient) condition for causality.
When developing a multiple regression model, there are various [assumptions](#sec-assumptionsRegression) that are important to evaluate.
In addition, it is important to pay attention for potential [multicollinearity](#sec-multipleRegressionMulticollinearity)—it may become difficult to detect a given [predictor variable](#sec-correlationalStudy) as [statistically significant](#sec-statisticalSignificance) due to the greater uncertainty around the parameter estimates.
